### 2011-01-04 doc/affiliates/greenery.tex

\documentclass{beamer}
\usetheme{Dresden}
\usepackage{enumerate,listings}

\setbeamertemplate{navigation symbols}{}
\begin{document}
\begin{frame}
\frametitle{libgreenery}
Goal: expose more memory references through independent contexts.
\begin{enumerate}
  \item Overhead between threads' references can't outweigh gain from prefetch
  \item Context switch must be as fast as possible ($<10$ns)
  \item Many threads must fit in same cache (tiny stacks)
  \item Dramatic limits on what you can run inside these
  \end{enumerate}
\begin{block}{API}
  Minimal as possible:
  \begin{enumerate}
  \item \texttt{thread\_spawn}, \texttt{thread\_yield}
  \item \texttt{thread\_block} (mostly avoid explicit synchronization)
  \item \texttt{prefetch\_and\_switch} (the key)
  \end{enumerate}
\end{block}
\end{frame}
\begin{frame}[fragile]
  \frametitle{Example usage}
Unlike hardware threads, we \emph{want} these to cooperate tightly.
\begin{lstlisting}
// shared
int examine_size, i = 0;
vertex examine[], next[];
// many threads run this:
while (i < examine_size) {
  vertex v = examine[i];
  vertex edges[] = prefetch_and_switch(edge_lists[v]);
  for (j = 0; j < edges.size; ++j) {
     vertex target = prefetch_and_switch(edges[j]);
     next.push_back(target);
  }
}
thread_block(barrier)
// swap queues, repeat
\end{lstlisting}
Rely mostly on lack of preemption for synchronization, work on shared
data, and prefetch/switch on any reference with poor locality.
\end{frame}
\end{document}