### 2012-07-18 doc/asplos13/implementation.tex
Rearranged implementation and made a pass on it.

\section{Implementation} \label{sec:implementation}

We now discuss implementation details for the major system components discussed in Section~\ref{sec:grappa}.

\subsection{Memory}

Grappa implements a software distributed shared memory. This memory is
exposed through two address spaces; one optimized for global access to
node-local data, and another optimized for high aggregate random
access bandwidth to low locality data.

\paragraph{2D global addresses:} All global addresses in Grappa are 64-bit values. Each byte of global
memory is associated with a {\em home core}; this core is responsible
for allocating and modifying that memory. Since today's commodity
processors support a 48-bit virtual address space \TODO{cite?}, Grappa
uses the spare bits to distinguish which address space a global
address belongs to, and to represent the home core of a global
address.

The first global address space is the {\em two-dimensional} address
space. This uses a traditional PGAS addressing model, where each
address is a tuple of a rank in the job (or global process ID) and an
address in that process. The lower 48 bits of the address hold a
virtual address in the process. The top bit is set to indicate that
this a 2D address. This leaves 15 bits for network endpoint ID, which
limits our scalability to $2^15$ endpoints.

Any local data can be made accessible by other nodes in the system by
wrapping the address and node ID into a 2D global address. This
address can then be accessed with delegate or cache operations. At the
destination the address is converted into a canonical x86 address by
replacing the upper bits with the sign-exteded upper bit of the
virtual address.

\paragraph{Linear global addresses:} The second global address space is the {\em linear} address space, so
named because nearby addresses can be treated as adjacent even if they
are stored on different nodes. Addresses in the linear address space
are partitioned across the cluster in a block-cyclic fashion. Choosing
the block size involves trading off sequential bandwidth against
aggregate random access bandwidth. Smaller block sizes help spread
data across all the memory controllers in the cluster, but larger
block sizes allow the locality-optimized memory controllers to provide
increased sequential bandwidth. The block size, which is configurable, is typically set to 64
bytes, or the size of a single hardware cache line, in order to
exploit spatial locality when available.

All linear addresses refer to data allocated from a global heap. The
heap metadata is stored on a single node. Currently all heap
operations serialize through this node; while this has been sufficient
for our benchmarks, parallel performance can be achieved through
combining \TODO{ cite MAMA and flat combining }. 2D addresses may
refer to memory allocated from a single processes' heap or from a
task's stack.

\TODO{Add a figure that depicts both forms of global addresses.}

All global memory operations are implemented with active
messages. Requestors construct a descriptor holding the current
task/worker ID and space for a response, issue their request active
message to the node given by the request address, and suspend
themselves. When the request arrives at the address' home node, the
operation is performed and a response active message is sent. When the
response arrives at the requesting node, the descriptor is filled in
with the results of the operation and the requesting task is woken.

\paragraph{Delegates:} Delegate operations are used for syncronization. Their modifications
are always executed at the home core of their address. Each task may
have only one outstanding delegate operation. We limit delegate
operations to operate on objects in the 2D address space or objects that
fit in a single block of the linear address space. No context switches
are allowed while the data it being modified. Given these
restrictions, we can ensure that delegate operations for the same address from multiple
requestors are always serialized through a single core in the system,
providing atomic semantics without using atomic operations.

Cache operations are used to exploit spatial or temporal
locality. They allow a single task to buffer a range of data, read or
modify it, and write the modified version back to its home node. All
acquires and releases are expressed explicitly by the
programmer. Cache operations may operate on data that spans multiple
blocks of the linear address space; the programmer provides the
starting address and length, and the cache issues as many active
messages as necessary to retrieve all necessary data.

\comment{If anything, this is just a footnote. -M

All current cache operations are incoherent; two tasks may
concurrently modify the same block with their writes interfereing
unless protected by synchronization through delegate operations. While
this has been sufficient for our benchmarks, it is possible to embed
synchronization in the cache active messages and reduce the number of
network round trips necessary.}

\subsection{Tasking}

\subsubsection{Tasks and workers}

Grappa tasks are 32-byte entities: a 64-bit function pointer plus
three 64-bit arguments. We use three arguments because tasks are
usually generated as part of a parallel loop decomposition, and thus
each task needs three kinds of data:
\begin{description}
\item[Function pointer] This 64-bit value indicates what routine
  should run. We configure the system nodes to disable address randomization, so that we can depend on the fact that function pointers are valid across process images.
\item[Private argument] This 64-bit value is typically used for task-specific
  data (e.g., a loop index).
\item[Shared argument] This 64-bit value is typically used for data shared
  between all tasks that are part of a loop.
\item[Synchronization argument] This 64-bit value is typically used to determine
  when all tasks that are part of a loop have finished. It is usually
  a global pointer to a synchronization object allocated at the core
  that spawned the task.
\end{description}
While these are the most common uses of the three task arguments, they
are treated as arbitray 64-bit values in the runtime, and can be used
for any purpose.

Tasks are not allocated any execution resources until the scheduler
decides to run them; when this occurs, tasks are matched with {\em
  workers}. Each worker is simply a collection of status bits and a
stack, allocated at a particular core.

\subsubsection{Context switching}

Context switching in Grappa is non-preemptive. Any Grappa routine that
does a potentially-long-latency operation calls a context-switch
routine that gives control of the core back to the scheduler. Users
may also explicity yield control to the scheduler.

As with other cooperative multithreading systems, we treat context
switches as function calls, saving and restoring only the callee-saved
state as specified in the x86-64 ABI \TODO{cite}. This involves saving
six general-purpose 64-bit registers and the stack pointer, as well as
the 16-bit x87 floating point control word and the SSE context/status
register. Thus, the minimum amount of state a cooperative context
switch routine must save according to the ABI is 62 bytes.

Since the compiler sees all calls to the context switch routine, we
can save even less state. Our context switch routine appears to the
compiler as inline assembly; we declare all the registers we need
to save as ``clobbered'' by the inline assembly routine, and the
compiler will issue its own save and restore code as needed. This allows the
compiler to avoid saving any registers that are not used, or are used
for temporary values that are not needed after the context switch.
\comment{I highly doubt this is important in real apps beyond our kernels.  -M

In
particular, only one of our applications does floating point
operations; all the other applications can safely ignore the x87 and
SSE2 registers as part of context switching.}

\subsubsection{Scheduling}

Each core in the Grappa system has its own independent scheduler. Each
scheduler has three main tasks to perform.  First, it must ensure that
its communication resources are serviced reasonably often. Second, it
must ensure that if a running task was waiting on a long-latency
operation and that operation has completed, the task will be
rescheduled. Third, if there are tasks waiting to be run and spare
execution resources, the tasks must be matched with idle workers.

Each scheduler has three queues:
\begin{description}
\item{\bf Ready worker queue} This is a FIFO queue of tasks that are
  matched with workers and are ready to execute.
\item{\bf Private task queue} This is a FIFO queue of tasks that must run on this core.
\item{\bf Public task queue} This is a LIFO queue of tasks that are
  waiting to be matched with workers. It is a local partition of a shared
  task pool; this is described in the next section.
\end{description}

\TODO{What is a worker?  We don't realyl describe this.  -M}

Whenever a task yields or suspends, the scheduler makes a decision
about what to do next. First, it determines if
the communication resources should be serviced. This is done on a
periodic basis. Second, it determines if any workers with running
tasks are ready to execute; if so, one is scheduled. Finally, if there
are no workers ready to run, but there are tasks waiting to be matched
with workers, an idle worker is woken (or a new worker is spawned),
matched with a task, and scheduled.

The scheduler uses its core's timestamp counter to track time. The
timestamp counter is read once per scheduling decision. Timestamp
values are not shared between cores. We use timestamp values only to
ensure periodic execution of communication maintenance tasks; being
wrong will only change the rate of these events with a potential
performance cost, but does not cause a correctness violation. Nevertheless, we
assume Grappa will run on machines with modern timestamp counter
implementations that are monotonic and frequency-compensated.

\subsubsection{Work stealing} \label{subsec:implementation-worksteal}
When a \checkme{grappa-process} runs
out of work, it becomes a ``thief'' and asks another other
\checkme{grappa-process} (``victim'') for work. If the victim has tasks
to spare in its public queue, they will be transferred to the thief's public queue.

Most study of work stealing assumes that a processor has one worker
thread and so it only steals when utilization would go to zero. 
Grappa, on the other hand, relies on having many worker threads per core. The
core would be fully-utilized if there is always something on the ready
queue. Even if there are many active tasks, if they are all suspended
for long-latency network requests, then the core is underutilized.
So, there is a choice: should the core use some of this idle time to
perform work stealing or should it just wait as it is possible that local tasks will
create more work. We choose to have a core inititiate a steal when all
of the following conditions hold: no workers are ready to run, 
the unstarted task queues are empty, and there are no outstanding steal requests. 
Having only one outstanding steal request throttles steals to
prevent flooding the network, but other scalable quieting mechanisms
would be possible, such as a voting tree\cite{scalableWorkStealingOrCilk98}
or lifelines \cite{lifelines}. Termination detection is not
built into the Grappa task scheduler, rather, it is considered a programming error for
the program to exit without syncing all tasks.

\TODO{where to put the following related}
Fork join computations: implement with futures or continuation passing
style
feed-forward computations: implement with scalable task counting (we
focus on this style in our benchmarks)
(breifly describe global task joiner)

choices we've made
-random victims
-steal from FIFO side of queue
-steal when at the moment readyQ empty and task queues empty
-steal min(chunk\_size, half)

Recent work has explored scalable work stealing in terms of efficient
termination detection.

\subsection{Communication}

In order to mitigate the low message injection rate limits of commodity
networks, Grappa's communication stack has two layers.

At the upper layer, Grappa implements asynchronous active messages
\TODO{cite vonEicken92}. Each message consists of a function pointer, an
optional argument payload, and an optional data payload. When a task
sends a message, the message is copied to a send queue associated with
the message's destination and the task continues execution.

Grappa's lower networking layer aggregates the upper layer's messages
to improve performance. In our applications we see that the bulk of
the messages sent are small. \TODO{include stats} Commodity networks,
even ones designed for high-performance computing, provide low
bandwidth utilization at small message sizes. \TODO{include stats?} To
make the best use of the network, we must convert our small messages
into large ones.

Grappa leverages its lightweight context switching to do this
conversion. When a task sends a message, the message is not
immediately sent. Instead, it is stored in a queue associated with its
destination, to be sent later. If the task requires a response, it
suspends itself and the scheduler finds other work to run while the
task is waiting. When the reply arrives, the suspended task is woken.

There are three situations in which a queue of aggregated messages is
sent. First, each queue has a message size threshold, chosen to give
reasonable network performance. If the size in bytes of a queue is
above the threshold, the contents of the queue are sent
immediately. Second, each queue has a wait time threshhold. If the
oldest message in a queue has been waiting longer than this threshold,
the contents of the queue are sent immediately, even if the queue size
is lower than the message size threshold.  Third, queues may be
explicitly flushed in situations where the programmer wants to
minimize the latency of a message at the cost of bandwidth
utilization.

The network layer is serviced by polling. Periodically when a context
switch occurs, the Grappa scheduler switches to the network polling
thread. This thread has three responsibilities. First, it polls the
lower-level network layer to ensure it makes progress. Second, it
deaggregates received messages and executes active message
handlers. Third, it checks to see if any aggregation queues have
messages that have been waiting longer than the threshold; if so, it
sends them.

Underneath the aggregation layer, Grappa uses the GASNet communication
library \TODO{cite GASNet} to actually move data. All interprocess
communication, whether on or off a cluster node, is handled by the
GASNet library. GASNet is able to take advantage of many communication
mechanisms, including Ethernet and Infiniband between nodes, as well as
shared memory within a node.

Some networks provide access to a remote machine's memory
directly. This would seem to be a good fit for a programming model
focused on global shared memory, but in fact we do not use it. RDMA
operations are subject to the same message rate limitations as all
other messages on these cards, and thus using raw RDMA operations for
our small messages would make inefficient use of bandwidth. Instead,
we implement remote memory operations with active messages. A byproduct of this design decision is that Grappa is not limited
to RDMA-capable networks.