### 2012-07-22 doc/asplos13/relatedwork.tex
some minor edits

\section{Background}

% The goal of the Grappa framework is to simplify the task of implementing
% irregular applications.
Grappa is built on many existing ideas in programming languages, systems and
architecture. In this section we discuss related frameworks and key enabling
technologies that Grappa builds upon.

\subsection{Comparative Frameworks}

Distributed graph processing frameworks like Pregel \cite{pregel:2010} and
Distributed GraphLab \cite{distgraphlab:vldb12} share similar goals. Pregel
adopts a bulk-synchronous parallel (BSP) execution model, which makes it
inefficient on workloads that could prioritize vertices. GraphLab, on the
other hand, schedules vertex computations individually, allowing
prioritization, which gives faster convergence in a variety of iterative
algorithms.\TODO{GraphLab is not general. Myers will expand on that here. Make
a strong point about generality.} Grappa also supports dynamic parallelism
with asynchronous execution, but parallelism is expressed as tasks or loop
iterations.

\subsection{Global memory}

Grappa implements a software distributed shared memory DSM system. Many
traditional software DSM systems are page based~\cite{Treadmarks,munin} and
aim to hide the fact that they are built in software from applications by
exploiting the processor's paging mechanisms. Instead Grappa, like other
partitioned global address space (PGAS) models, implements its DSM at the
language, rather than system level. Languages such as Chapel \cite{Chapel},
X10 \cite{X10}, and UPC \cite{UPC} make accesses to shared structures look
like normal memory references. As we describe later, Grappa chooses a middle
ground, where global addresses are explicit in the API and local accesses are
emitted conventionally by the compiler. Similar to DSM systems Grappa provides
a caching and coherence model to ensure high performance and consistent
results. Cache operations are exposed to in the API, however, allowing
software to optimize their usage.

\subsection{Multithreading}

Grappa uses multithreading to tolerate memory latency. This is a well known
technique. Hardware implementations include the Tera MTA \cite{Tera}, Cray XMT
\cite{XMT}, Simultaneous multithreading \cite{SMT}, MIT Alewife
\cite{Alewife}, Cyclops \cite{Cyclops}, and even GPUs \cite{fatahalian}. As we
describe in this paper, Grappa use a lightweight user-mode task scheduler to
multiplex thousands of tasks on a single processing core. The large number of
tasks is required because of the extremely large internode latency.