### 2012-07-22 doc/asplos13/grappa.tex
simplifying intro

\section{Grappa} \label{sec:grappa}

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.95\columnwidth]{figs/system-overview}
\begin{minipage}{0.95\columnwidth}
  \caption{\label{fig:grappa} Grappa system overview}
\end{minipage}
\vspace{-3ex}
\end{center}
\end{figure}

From the perspective of a software developer, Grappa
(Figure~\ref{fig:grappa}) provides two main software components: a
\textbf{tasking system} with lightweight multithreading to tolerate
latency and workstealing to automate load balance; and a
\textbf{distributed shared global memory} providing high aggregate
random access bandwidth for both normal and synchronized operations. We
now discuss the programmer's view of Grappa's main capabilities,
including a subset of the Grappa API. In Section~\ref{implementation},
we discuss the details of their implementation as well as a third
component, the network communication layer that is key to performance
but largely invisible to software developers.

\TODO{Should we have a code example here, and walk thru it as we explain the
components and parts of the API?}

\subsection{Tasks}

%The basic unit of locality in Grappa is the {\em core}. Each core is
%responsible for a section of the global memory

The basic unit of execution in Grappa is a {\em task}. Tasks are
small; just a function pointer and arguments. Tasks are only
queued when they are spawned; later, when resources are free, they are
allocated a stack, bound to a core, and executed.

Grappa's tasks give up control of their core when when they perform
long-latency operations, allowing the processor to remain busy while
waiting for the operation to complete. This is most often done
implicitly inside a call into Grappa's API, but it can also be done
explicitly by the programmer using the calls shown in
Figure~\ref{fig:scheduling}. \TODO{cite UPC split phase reads and
writes, but include that we provide a way to overlap computation}

All Grappa scheduling is done in userspace, and minimal state is saved,
in order to minimize the context switch time. In our experiments, we saw
context switch times as low as \checkme{40ns}. This lightweight context
switching is Grappa's key enabling feature. It gives us the ability to
tolerate latency. Given that ability, we are able to \textbf{trade
latency for throughput}: by {\em increasing} latency in key components
in the system we are able to increase our aggregate random access
bandwidth, our synchronization bandwidth, and our ability to tolerate
load.

\begin{figure}[htbp]
  \begin{center}
    \begin{description}\small
    \item[ \texttt{ yield() } ] \hfill \\
      Gives up control of core to scheduler, queuing task to be scheduled again soon
    \item[ \texttt{ suspend() } ] \hfill \\
      Gives up control of core to scheduler
    \item[ \texttt{ wake( task * $t$ ) } ] \hfill \\
      Enqueues $t$ to be scheduled again soon
    \end{description}
    \begin{minipage}{0.95\columnwidth}
      \caption{\label{fig:scheduling} Grappa API: scheduling} %{-4ex}}
    \end{minipage}
    %\vspace{-3ex}
  \end{center}
\end{figure}

\subsection{Expressing parallelism}

A programmer's goal in coding with Grappa should be to express as much
parallelism as possible without worrying about where it will execute.
Grappa then chooses where and when to exploit this parallelism,
scheduling as much work as necessary to tolerate network latencies and
keep cores busy.

Grappa provides four methods for expressing parallelism, shown in
Figure~\ref{fig:expressing-parallelism}. The first is explicit task
spawns. When the programmer identifies work that can be done in
parallel, the work may be wrapped up in a function and queued with its
arguments for later execution using a \texttt{spawn}. If the
programmer wishes to spawn a task on a specific core in the system,
or at the home core of a particular memory location, Grappa provides a
\texttt{spawn\_on} call for this purpose.

The next method for expressing parallelism is a parallel for loop, where
the number of iterations must be known at loop entry. The programmer
specifies a function pointer along with start and end indices and an
optional threshold to control parallel overhead. Grappa does {\em
recursive decomposition} of iterations, similar to Cilk's cilk\_for
construct~\cite {cilkforimplementation} \comment{could only find slides
page6 of
http://www.clear.rice.edu/comp422/lecture-notes/comp422-2012-Lecture5-
Cilk++.pdf    ---- Good enough.  Cite it.  -Mark}.  It generates a
logarithmically-deep tree of tasks, stopping to execute the loop body
when the number of iterations is below the required threshold. Finally,
a programmer may want to run a small piece of code on a particular core
in the system without waiting for execution resources to be available.
Grappa provides the \texttt{call\_on} call for this purpose.

\begin{figure}[htbp]
  \begin{center}
    \begin{description}\small
    \item[ \texttt{spawn( void (*fp)(args) )} ] \hfill \\
      Creates a new stealable task
    \item[ \texttt{spawn\_on( core, (*fp)(args) )} ] \hfill \\
      Creates a new private task that will run on a specific core 
    \item[ \texttt{parallel\_for( (*fp)(args), start, end )} ] \hfill \\
      Executes iterations of a loop as stealable tasks 
    \item[ \texttt{call\_on( core, (*fp)(args) )} ] \hfill \\ 
      Runs a limited function on a specific core without consuming
      Grappa execution resources 
    \end{description}
    \begin{minipage}{0.95\columnwidth}
      \caption{\label{fig:expressing-parallelism} Grappa API: expressing parallelism} % \vspace{-4ex}}
    \end{minipage}
    %\vspace{-3ex}
  \end{center}
\end{figure}

\subsection{Memory}

Applications written for Grappa utilize two forms of memory: local and
global.

Local memory is local to a single core in the system.  Accesses occur
through conventional pointers.  The compiler emits an access and the
memory is manipulated directly.  Applications use local accesses for a
number of things in Grappa: the stack associated with a task, accesses
to localized global memory in caches (see below), and accesses to
debugging infrastructure that is local to each system node.  Local
pointers cannot access memory on other cores, and are valid only on
their home core.

Large data that is expected to be shared and accessed with low locality is
stored in Grappa's global memory. All global data must be accessed through
calls into Grappa's API, shown in Figure~\ref{fig:accessing-memory}.

\paragraph{Global memory allocation}
Grappa provides two methods for {\emph storing} data in the global memory. The
first is a distributed heap striped across all the machines in the
system. The \texttt{global\_malloc} and \texttt{global\_free} calls
are used to allocate and deallocate memory in the global heap; on
allocation, a global pointer is returned. Grappa also allows any local
data on a core's stacks or heap to be exported to the global address
space to be made accessible to other cores across the system.

\paragraph{Global memory access} There are two approaches to {\emph
accessing} global memory. When the programmer expects a computation on
shared data to have spatial locality to exploit, {\em cache} operations
may be used. When there is no locality to exploit, {\em delegate}
operations are used.

\textbf{Explicit caching.} Grappa provides an API to fetch a global
pointer of any length and return a local pointer to a cached copy of the
global memory.  Grappa cache operations have the usual read-only and
read-write variants, along with a write-only variant used to initialize
data structures. Languages for distributed shared memory systems have
done optimizations to achieve a similar goal. For example, the UPC
compiler coalesces struct and array accesses into remote get/put
\cite{Chen:2005}, and Fortran D compiler's message vectorization hoists
small messages out of a loop \cite{FortranD:1992}. Caching in Grappa
additionally provides a mechanism for exploiting temporal locality by
operating on the data locally.

\textbf{Delegate operations.}
When the access pattern has low-locality, it is more efficient to modify
the data on its home core rather than bringing a copy to the requesting
core and returning it after modification. Delegate operations provide
this capability. Applications can dispatch computation to be performed
on individual machine-word sized chunks of global memory to the memory
system itself (e.g., \emph{fetch-and-add}).  Delegate operations,
proposed in \cite{Nelson:hotpar11} and \cite{delegated:oopsla11}, are also the primary synchronization method in Grappa.

\begin{figure}[htbp]
  \begin{center}
    \begin{description}\small
      \item[ \texttt{ global\_address global\_malloc( size )} ] \hfill \\
      \item[ \texttt{ global\_free( global\_address )} ] \hfill \\
        Allocates and frees memory in the global heap
      \item[ \texttt{ delegate\_read( global\_address, local\_var )} ] 
      \item[ \texttt{ delegate\_write( global\_address, local\_var )} ] %\vspace{-2ex}
      \item[ \texttt{ delegate\_cas( global\_address, local\_var )} ] %\vspace{-2ex}
      \item[ \texttt{ delegate\_inc( global\_address, local\_var )} ] %\vspace{-2ex} 
\hfill \\
        Performs a memory operation at the home core of a global address
      \item[ \texttt{ cache\_acquire( global\_address, local\_buf, \{RO,RW,WO\})} ]
      \item[ \texttt{ cache\_release( global\_address, local\_buf )} ] %\vspace{-2ex} 
\hfill \\
        Perform cache operations \TODO{expand}
    \end{description}
    \begin{minipage}{0.95\columnwidth}
      \caption{\label{fig:accessing-memory} Grappa API: accessing memory} %{-4ex}}
    \end{minipage}
    %\vspace{-3ex}
  \end{center}
\end{figure}