### 2012-07-23 doc/asplos13/paper.tex
formatting job

\input{header}

\begin{document}

\title{Grappa: Scalable Computing on Mass-Market Hardware for Irregular, Low-Locality Applications}
%\title{All Your Graphs Belong to Us}
\authorinfo{\vspace{-2ex}Anonymous for review\vspace{-6ex}}{}{}

\maketitle
\begin{abstract}
\input{abstract}
\end{abstract}

\TODO{Checklist}\\
\TODO{pages<=12 for everything}\\
\TODO{ACM format, 9pt font --- done}\\
\TODO{include page numbers --- done}\\
\TODO{readable with black and white}\\
\TODO{no author names, affiliations, personal acknowledgments, and any
    other hints of identity --- done}\\
\TODO{conceal authorship of prior work}\\

\input{intro}

\input{relatedwork}

\input{grappa}
\input{implementation}

\input{method}

\input{evaluation}

\input{acknowledgements}

\subsection{References}
\bibliographystyle{abbrv}
\bibliography{paper}

\end{document}