### 2012-07-23 doc/asplos13/communication.tex
Super redux

\section{Communication}

In order to mitigate the low message injection rate limits of commodity
networks, Grappa's communication stack has two layers: one for
user-level messages and one for network-level messages.

At the upper layer, Grappa implements asynchronous active messages
\cite{vonEicken92}. Each message consists of a function pointer, an
optional argument payload, and an optional data payload. When a task
sends a message, the message is copied to a send queue associated with
the message's destination and the task continues execution.

Grappa's lower networking layer aggregates the upper layer's messages
to improve performance. Commodity networks including infiniband
achieves their peak bisection bandwidth \emph{only} when the packet
sizes are relatively large---on the order of multiple kilobytes. The
reason for this discrepancy is the combination of overheads associated
with handling each packet (in terms of bytes that form the actual
packet, processing time at the card and processing on the
CPU within the driver stack). Our measurements confirm manufacturers
published data~\cite{infinibandbandwidth}, that with this packet size
the bisection bandwidth is only a small fraction, less than
~\checkme{5\%} of the peak bisection bandwidth.

In our experiments the vast majority of requests were smaller than
\checkme{255} bytes \TODO{more precise? median?}, far too small to make
efficient use of the network. To make the best use of the network, we
must convert our small messages into large ones
(Section~\ref{sec:results-aggregation}).  When a task sends a message,
it is not immediately sent, but rather placed in a queue specific to the
destination.

There are three situations in which a queue of aggregated messages is
sent. First, each queue has a message size threshold, chosen to give
reasonable network performance (\checkme{2048 bytes}). If the size in
bytes of a queue is above the threshold, the contents of the queue are
sent immediately. Second, each queue has a wait time threshold
(\checkme{200ms}). If the oldest message in a queue has been waiting
longer than this threshold, the contents of the queue are sent
immediately, even if the queue size is lower than the message size
threshold.  Third, queues may be explicitly flushed in situations where
the programmer wants to minimize the latency of a message at the cost of
bandwidth utilization.

The network layer is serviced by polling. Periodically when a context
switch occurs, the Grappa scheduler switches to the network polling
thread. This thread has three responsibilities. First, it polls the
lower-level network layer to ensure it makes progress. Second, it
deaggregates received messages and executes active message
handlers. Third, it checks to see if any aggregation queues have
messages that have been waiting longer than the threshold; if so, it
sends them.

Underneath the aggregation layer, Grappa uses the \gasnet~communication
library~\cite{gasnet} to actually move data. All interprocess
communication, whether on or off a cluster node, is handled by the
\gasnet~library. \gasnet~is able to take advantage of many communication
mechanisms, including ethernet and infiniband between nodes, as well as
shared memory within a node.

Some networks provide access to a remote machine's memory directly. This
would seem to be a good fit for a programming model focused on global
shared memory, but in fact we do not use it. In our experiments, we
found that RDMA operations are subject to the same message rate
limitations as all other messages on these cards, and thus using raw
RDMA operations for our small messages would make inefficient use of
bandwidth. Instead, we implement remote memory operations with active
messages. A byproduct of this design decision is that Grappa is not
limited to RDMA-capable networks.