### 2012-07-23 doc/asplos13/memory.tex

\section{Memory}

Applications written for Grappa utilize two forms of memory: local and
global.

Local memory is local to a single core in the system.  Accesses occur
through conventional pointers.  The compiler emits an access and the
memory is manipulated directly.  Applications use local accesses for a
number of things in Grappa: the stack associated with a task, accesses
to localized global memory in caches (see below), and accesses to
debugging infrastructure that is local to each system node.  Local
pointers cannot access memory on other cores, and are valid only on
their home core.

Large data that is expected to be shared and accessed with low locality is
stored in Grappa's global memory. All global data must be accessed through
calls into Grappa's API, shown in Figure~\ref{fig:accessing-memory}.

\paragraph{Global memory allocation}
Grappa provides two methods for {\emph storing} data in the global memory. The
first is a distributed heap striped across all the machines in the
system. The \texttt{global\_malloc} and \texttt{global\_free} calls
are used to allocate and deallocate memory in the global heap; on
allocation, a global pointer is returned. Grappa also allows any local
data on a core's stacks or heap to be exported to the global address
space to be made accessible to other cores across the system.

\begin{figure}[htbp]
  \begin{center}
    \begin{description}\small
      \item[ \texttt{ global\_address global\_malloc( size )} ] \hfill \\
      \item[ \texttt{ global\_free( global\_address )} ] \hfill \\
        Allocates and frees memory in the global heap
      \item[ \texttt{ delegate\_read( global\_address, local\_var )} ] 
      \item[ \texttt{ delegate\_write( global\_address, local\_var )} ] %\vspace{-2ex}
      \item[ \texttt{ delegate\_cas( global\_address, local\_var )} ] %\vspace{-2ex}
      \item[ \texttt{ delegate\_fetch\_inc( global\_address, local\_var )} ] %\vspace{-2ex} 
\hfill \\
        Performs a memory operation at the home core of a global address
      \item[ \texttt{ cache\_acquire( global\_address, local\_buf, \{RO,RW,WO\})} ]
      \item[ \texttt{ cache\_release( global\_address, local\_buf )} ] %\vspace{-2ex} 
\hfill \\
        Perform cache operations \TODO{expand}
    \end{description}
    \begin{minipage}{0.95\columnwidth}
      \caption{\label{fig:accessing-memory} Grappa API: accessing memory} %{-4ex}}
    \end{minipage}
    %\vspace{-3ex}
  \end{center}
\end{figure}

\paragraph{Global memory access} There are two approaches to {\emph
accessing} global memory. When the programmer expects a computation on
shared data to have spatial locality to exploit, {\em cache} operations
may be used. When there is no locality to exploit, {\em delegate}
operations are used.

All global addresses in Grappa are 64-bit values. Each byte of global
memory is associated with a {\em home core}; this core is responsible
for allocating and modifying that memory. Since today's commodity
processors support a 48-bit virtual address space \cite{AMD64}, Grappa
uses the spare bits to distinguish which address space a global address
belongs to, and to represent the home core of a global address.

Global memory in Grappa is exposed through two distinct, non-overlapping
address spaces; \textbf{2D global addresses}, optimized for global
access to node-local data, and \textbf{linear global addresses},
optimized for high aggregate random access bandwidth to low locality
data.

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.95\columnwidth]{figs/memory-structure}
\begin{minipage}{0.95\columnwidth}
  \caption{\label{fig:memory-structure} Grappa memory structure}
\end{minipage}
\vspace{-3ex}
\end{center}
\end{figure}

\paragraph{2D global addresses} The first global address space is the {\em
two-dimensional} address space. This uses a traditional PGAS addressing model,
where each address is a tuple of a rank in the job (or global process ID) and
an address in that process. The lower 48 bits of the address hold a virtual
address in the process. The top bit is set to indicates that the reference is
a 2D address. This leaves 15 bits for network endpoint ID, which limits our
scalability to $2^{15}$ endpoints.

Any node-local data can be made accessible by other nodes in the system
by wrapping the address and node ID into a 2D global address. This
address can then be accessed with a delegate and can also be cached by
other nodes. At the destination the address is converted into a
canonical x86 address by replacing the upper bits with the sign-extended
upper bit of the virtual address. 2D addresses may refer to memory
allocated from a single processes' heap or from a task's stack.
Figure~\ref{fig:memory-structure} shows how 2D and linear addresses can
refer to other cores' memory.

\paragraph{Linear global addresses} The second global address space is
the {\em linear} address space, so named because nearby addresses can be
treated as adjacent even if they are stored on different nodes.
Addresses in the linear address space are partitioned across the cluster
in a block-cyclic fashion. Choosing the block size involves trading off
sequential bandwidth against aggregate random access bandwidth. Smaller
block sizes help spread data across all the memory controllers in the
cluster, but larger block sizes allow the locality-optimized memory
controllers to provide increased sequential bandwidth. The block size,
which is configurable, is typically set to 64 bytes, or the size of a
single hardware cache line, in order to exploit spatial locality when
available.

All linear addresses refer to data allocated from a global heap. The
heap metadata is stored on a single node. Currently all heap operations
serialize through this node; while this has been sufficient for our
benchmarks, in the future Grappa will provide parallel performance
through combining~\cite{MAMA,flatcombining}.

\TODO{Add a figure that depicts both forms of global addresses.}

All global memory operations are implemented with active messages.
Requesters construct a descriptor holding the current task ID and space
for a response, issue their request active message to the node given by
the request address, and suspend themselves. When the request arrives at
the address' home node, the operation is performed and a response active
message is sent. When the response arrives at the requesting node, the
descriptor is filled in with the results of the operation and the
requesting task is placed back on the ready queue.

\textbf{Explicit caching.} Grappa provides an API to fetch a global
pointer of any length and return a local pointer to a cached copy of the
global memory.  Grappa cache operations have the usual read-only and
read-write variants, along with a write-only variant used to initialize
data structures. Languages for distributed shared memory systems have
done optimizations to achieve a similar goal. For example, the UPC
compiler coalesces struct and array accesses into remote get/put
\cite{Chen:2005}, and Fortran D compiler's message vectorization hoists
small messages out of a loop \cite{FortranD:1992}. Caching in Grappa
additionally provides a mechanism for exploiting temporal locality by
operating on the data locally.

Under the hood, Grappa performs the mechanics of gathering chunks of
data from multiple system nodes and presenting a conventional appearing
linear block of memory as a local pointer into a cache. The strategy
employed is to issue all the constituent requests of a cache access
request and then yield until all responses have occurred.  Currently,
Grappa caches are \emph{not} coherent, requiring the programmer to
maintain consistent access to data.  Future work will develop a software
directory based coherence scheme to simplify consistent access to global
data.

\textbf{Delegate operations.}
When the access pattern has low-locality, it is more efficient to modify
the data on its home core rather than bringing a copy to the requesting
core and returning it after modification. Delegate operations provide
this capability. Applications can dispatch computation to be performed
on individual machine-word sized chunks of global memory to the memory
system itself (e.g., \emph{fetch-and-add}).  Delegate operations,
proposed in \cite{Nelson:hotpar11} and \cite{delegated:oopsla11}, are also the primary synchronization method in Grappa.

Delegate operations are always executed at the home core of their
address, and while arbitrary memory operations can be delegated, we
restrict the use of delegate operations in three ways to make them more
useful for synchronization. First, we limit each task to one outstanding
delegate operation to avoid the possibility of reordering in the
network. Second, we limit delegate operations to operate on objects in
the 2D address space or objects that fit in a single block of the linear
address space so they can be satisfied with a single network request.
Finally, no context switches are allowed while the data is being
modified. Given these restrictions, we can ensure that delegate
operations for the same address from multiple requesters are always
serialized through a single core in the system, providing atomic
semantics without using atomic operations.

\paragraph{Example}

\begin{figure*}[htb]
\begin{center}
  \includegraphics[width=1.5\columnwidth]{figs/delegate-cache}
\begin{minipage}{1.5\columnwidth}
  \caption{\label{fig:delegate-cache} Delegation and cache example.}
\end{minipage}
\vspace{-3ex}
\end{center}
\end{figure*}

Figure~\ref{fig:delegate-cache} shows an example of delegate and cache
operations. In step 1, a core allocates an array in the global
heap. It then spawns two tasks on remote cores to increment elements
of the array.

The first task increments two elements of the array using cache
operations. In step 2, the task is invoked. A cache request is issued
for two adjacent integers starting at the second element of the
array. Since these element are stored in the memories of two different
cores, this requires the sending of two messages in step 3. The task
is suspended until both responses arrive in step 4. The data carried
in these responses is stored in the local buffer. The elements are
then increment in the buffer. Then in step 5, the modified data is
sent back to the home node. Acknoledgements are returned in step 6 so
the task knows when the writes are complete.

The second task increments an element of the array with a delegate
operation. In step 7, the task is invoked. A delegate request is sent
to the home core of the array element with the increment value. The
task suspends until the response is received. In step 8, the increment
is executed on the remote core. A response is returned in step 9 with
the previous value of the array element.