### 2012-07-23 doc/asplos13/relatedwork.tex
another pass on related work. still needs generality claims

\section{Background}

% The goal of the Grappa framework is to simplify the task of implementing
% irregular applications.
Grappa is built on many existing ideas in programming languages, systems and
architecture. In this section we discuss related frameworks and key enabling
technologies that Grappa builds upon.

\paragraph{Comparable frameworks} Distributed graph processing frameworks like
Pregel \cite{pregel:2010} and Distributed GraphLab \cite{distgraphlab:vldb12}
share similar goals as Grappa. Pregel adopts a bulk-synchronous parallel (BSP)
execution model, which makes it inefficient on workloads that could prioritize
vertices. GraphLab, on the other hand, schedules vertex computations
individually, allowing prioritization, which gives faster convergence in a
variety of iterative algorithms.\TODO{GraphLab is not general. Myers will
expand on that here. Make a strong point about generality.} Grappa also
supports dynamic parallelism with asynchronous execution, but parallelism is
expressed as tasks or loop iterations.\TODO{Should RAMcloud go here?}.

\paragraph{Global memory} Grappa includes a custom implementation of a
software distributed shared memory (DSM) system. Many traditional software DSM
systems are page based~\cite{Treadmarks,munin} and aim to hide the fact that
they are built in software from applications by exploiting the processor's
paging mechanisms, therefore relying heavily on locality. Instead, Grappa,
like other partitioned global address space (PGAS) models, implements its DSM
at the language, rather than system level. Languages such as Chapel
\cite{Chapel}, X10 \cite{X10}, and UPC \cite{UPC} make accesses to shared
structures look like normal memory references. As we describe later, Grappa
chooses a middle ground, where global addresses are explicit in the API and
local accesses are emitted conventionally by the compiler. Similar to DSM
systems, Grappa provides a consistency model that ensure high performance and
with good programmability properties. % Caching is Cache operations are explicit in the  exposed to in the
% API, however, allowing software to optimize their usage.

We decided to build a custom distributed shared memory implementation, as
opposed to using an existing implementation, for two key reasons: (1) in order
to support the large number of tasks required for latency tolerance, the
implementation of the global memory system is tightly coupled to the task
scheduler; and (2) since commodity networks only do efficient RDMA or ordinary
transmission of small messages, the implementation of the global memory system
must make use of the network message aggregator.

%

\paragraph{Multithreading}
Grappa uses multithreading to tolerate memory latency. This is a well known
technique. Hardware implementations include the Tera MTA \cite{Tera}, Cray XMT
\cite{XMT}, Simultaneous multithreading \cite{SMT}, MIT Alewife
\cite{Alewife}, Cyclops \cite{Cyclops}, and even GPUs \cite{fatahalian}. As we
describe in this paper, Grappa use a lightweight user-mode task scheduler to
multiplex thousands of tasks on a single processing core. The large number of
tasks is required because of the extremely large internode latency.