### 2012-07-23 doc/asplos13/paper.tex
minor stuff

\input{header}

\begin{document}

\title{Grappa: Scalable Computing on Commodity Hardware \\for Irregular Applications}
%\title{All Your Graphs Belong to Us}
\authorinfo{\vspace{-2ex}Anonymous for review\vspace{-6ex}}{}{}

\maketitle
\begin{abstract}
\input{abstract}
\end{abstract}

\input{intro}

\input{relatedwork}

\input{overview}

\input{tasks}

\input{memory}

\input{communication}

\input{method}

\input{evaluation}

\input{conclude}

%\input{acknowledgements}

\subsection{References}
\bibliographystyle{abbrv}
\bibliography{paper}

\end{document}