### 2012-07-23 doc/asplos13/relatedwork.tex
Minor spacing fixes (Preston)

\section{Background}

Grappa is built on many existing ideas in programming languages, systems and
architecture. In this section we discuss related frameworks and key enabling
technologies that Grappa builds upon.

\paragraph{Comparable frameworks} Distributed graph processing
frameworks like Pregel~\cite{pregel:2010} and Distributed
GraphLab~\cite{distgraphlab:vldb12} share similar goals as
Grappa. Pregel adopts a bulk-synchronous parallel (BSP) execution
model, which makes it inefficient on workloads that could prioritize
vertices. GraphLab, on the other hand, schedules vertex computations
individually, allowing prioritization, which gives faster convergence
in a variety of iterative algorithms.  GraphLab, however, imposes a
rigid computation model where programmers must express computation as
transformations on a vertex and its edge list only, with only
information from adjacent vertexes. Pregel is only slightly less
restrictive, as the input data can be any vertex in the graph.  Grappa
also supports dynamic parallelism with asynchronous execution, but
parallelism is expressed as tasks or loop iterations, which is a far
more general programming model for irregular computation tasks.

\paragraph{Global memory} Grappa includes a custom implementation of a
software distributed shared memory (DSM) system. Many traditional
software DSM systems are page based~\cite{Treadmarks,munin} and aim to
hide the fact that they are built in software from applications by
exploiting the processor's paging mechanisms, therefore relying
heavily on locality. Instead, Grappa, like other partitioned global
address space (PGAS) models, implements its DSM at the language,
rather than system level. Languages such as
Chapel~\cite{Chamberlain:2007}, X10~\cite{X10:2005}, and
UPC~\cite{upc:2005} make accesses to shared structures look like
normal memory references. As we describe later, Grappa chooses a
middle ground, where global addresses are explicit in the API and
local accesses are emitted conventionally by the compiler.  While
Grappa's DSM system is conceptually similar to prior work, its
implementation is tuned for irregular computations.  Past DSM work,
being page-based, could exploit the RDMA capabilities of network
hardware to move large page-sized blocks of data from node to node.
In our experience, when these networks move small blocks of data (a
few bytes), only a fraction of the available bandwidth is achieved.
In addition, the DSM system in Grappa ends up being tightly coupled to
the task scheduler in order to overlap long latency memory operations
with useful computation.  For these two reasons it became necessary to
build a new DSM system specifically for Grappa.

%% We could be stronger here on why Grappa's task library is different,
%% but we are running out of time.  -Mark

\paragraph{Multithreading} Grappa uses multithreading to tolerate
memory latency. This is a well known technique. Hardware
implementations include the Tera MTA~\cite{tera:mta1}, Cray
XMT~\cite{feo:xmt}, Simultaneous multithreading~\cite{tullsen:smt},
MIT Alewife~\cite{agarwal:alewife}, Cyclops~\cite{almasi:cyclops}, and
even GPUs~\cite{gpus}. As we describe in this paper, Grappa use a
lightweight user-mode task scheduler to multiplex \emph{thousands\/} of
tasks on a single processing core. The large number of tasks is
required because of the extremely high internode latency Grappa is
mitigating.  Grappa's task library employs several optimizations: an
extremely fast task switch, a small task size, and judicious use of
hardware prefetching to bring task state into the cache long before
that task is actually scheduled.