### 2012-10-16 doc/asplos13/rebuttal.txt
myers first quick rebuttal pass

We thank our reviewers for their thoughtful comments.

Some reviewers (1, 4, 5) did not see a novel contribution in our work. We
believe that our system is the first runtime designed specifically to tolerate
latency by switching between thousands of concurrent tasks per core. We are
able to keep cores busy while performing remote memory, synchronization, and
work-stealing operations. We aggregate small shared memory references into
larger network messages to better utilize the network. While the individual
ideas used to build Grappa are generally well-known, we know of no other
runtime that combines these ideas to trade latency for throughput.

Reviewer 3 takes issue with our characterization of our performance as "on
par" with that of the XMT. We are able to provide a simple, general
programming model similar to that of the XMT on already-installed commodity
clusters with at most a 2.5x slowdown, even though the XMT has a 10x faster
network injection rate and 20x faster context switches than we do. We believe
this slowdown is within the realm of optimization of the runtime features
already covered in the paper---improvements to our
network layer should allow us to double the number of cores per node we use,
and improvements to our scheduler should allow us to increase the rate at
which each core injects messages into the network by ~30 percent.  [We should
be precise about "similar programming model" to avoid more overstated claims,
because in Grappa's current form it is not quite the same. The main difference
is not a difficult one but it exists: in our benchmarks we take some locality into account: e.g. caching edgelist chunks]

Reviewer 2 asks about algorithmic modifications such as Beamer's BFS
transformation. Both the XMT and Grappa can benefit from such
transformations. Beamer's transformation increases BFS performance in
two ways: first, by reducing the number of edges traversed, and
second, by reducing the number of synchronizing operations
executed. Both the XMT and Grappa would benefit from the reduction in
edges traversed, but since Grappa's remote references are relatively
more costly than the XMT's, we would expect Grappa to benefit
more. Atomic operations on both platforms are no more expensive than
normal memory operations, so we would expect no significant benefit
from the reduction in synchronizing operations. Many transformations
also benefit by exploiting locality: Grappa provides mechanisms to do
so easily, while the hashed addressing of the XMT makes exploiting
locality difficult.

Reviewer 1 suggests comparing Grappa against other benchmark
implementations on commodity clusters in order to evaluate the benefit
of the programming model. In this work, we focused on providing a
similar programming model to the XMT, which is known to be simple
relative to distributed memory machines. We will explore programmer
productivity and compare against other commodity-cluster-based
implementations in future work.

Our evaluation was written in haste; in a final version we would discuss key
metrics including scheduling overhead and aggregator performance limitations
<Reviwer 2>. The evaluation of the scheduler would break down scheduling
overhead into checking the polling thread, checking other
queues, doing load balancing, and context switch time, and compare the average overhead to the time a task executes between yields.

We thank Reviewer 4 for expanding our understanding of the related
work. We will incorporate the suggestions.

[scalability?]

[citation for current memory hierarchy != irregular apps?]