### 2012-10-16 doc/asplos13/rebuttal.txt
some edits, reformatting

We thank the reviewers for their thoughtful comments.

>>> What is the novelty of the work? (Reviewers 1, 4, 5)

We believe that our system is the first runtime designed specifically to
tolerate latency by switching between thousands of concurrent tasks per core.
We are able to keep cores busy while performing remote memory,
synchronization, and work-stealing operations. We aggregate messages from
unrelated threads into larger network messages to better utilize the network.
While the individual ideas used to build Grappa are generally well-known, we
know of no other runtime that combines these ideas to trade latency for
throughput to build a general-purposed parallel runtime for large-scale applications.

>>> Why a performance comparison with XMT? (Reviewer 3)

We are able to provide a simple, general programming model similar to that of
the XMT on ubiquitous commodity clusters with at most a 2.5x slowdown (and at
best TODO faster), even though the XMT has a 10x faster network injection rate
and 20x faster context switches than we do. We believe this slowdown is within
the realm of optimization of the runtime features already covered in the
paper---improvements to our network layer should allow us to double the number
of cores per node we use, and improvements to our scheduler should allow us to
increase the rate at which each core injects messages into the network by ~30
percent.

[This would just complicate things. I would leave it out... We should be
precise about "similar programming model" to avoid more overstated claims,
because in Grappa's current form it is not quite the same. The main difference
is not a difficult one but it exists: in our benchmarks we take some locality
into account: e.g. caching edgelist chunks]

>>> Why not compare with other benchmarks that running commodity clusters?
(Reviewer 1)

In this work, we focused on providing a similar programming model to the XMT,
which is known to be simple relative to distributed memory machines. We will
explore programmer productivity and compare against other
commodity-cluster-based implementations in future work.

[Can we say something more about how we would compare based on published results?]

>>> Evaluation section is unpolished and incomplete. (Reviewer 2)

We will cleanup and expand our evaluation. In the final version we will
discuss key metrics including scheduling overhead and aggregator performance
limitations. The evaluation of the scheduler would break down scheduling
overhead into checking the polling thread, checking other queues, doing load
balancing, and context switch time, and compare the average overhead to the
time a task executes between yields. We will expand the evaluation with a detailed study of where time goes, since Grappa has a lightweight tracing infrastructure --- we did not add this to the evaluation due to lack of time.

[You can explain all these points in the rebuttal. You just can't add more benchamrks or major results. TODO: what we can say in rebuttal without more data: 1. Why BC and BFS are
worse while UTS is better than XMT. I think the story is that our injection
rate is not as good and UTS wins because of XMT not scaling 2. Why scaling of
apps is BC < BFS < UTS.]

>>> What would be the impact of better algorithms (e.g., Beamer's BFS) in the
performance comparisons? (Reviewer 2)

Both the XMT and Grappa can benefit from such transformations. Beamer's
transformation increases BFS performance in two ways: first, by reducing the
number of edges traversed, and second, by reducing the number of synchronizing
operations executed. Both the XMT and Grappa would benefit from the reduction
in edges traversed, but since Grappa's remote references are relatively more
costly than the XMT's, we would expect Grappa to benefit more. Atomic
operations on both platforms are no more expensive than normal memory
operations, so we would expect no significant benefit from the reduction in
synchronizing operations. Many transformations also benefit by exploiting
locality: Grappa provides mechanisms to do so easily, while the hashed
addressing of the XMT makes exploiting locality difficult.

>>> Missing related work. (Reviewer 4)

Thanks for pointing to extra related work, we will add these to the final version.

[scalability?]

[citation for current memory hierarchy != irregular apps?]