### 2013-01-29 doc/xps2013/header.tex
XPS 2013 rough in

\documentclass[11pt]{article}

\usepackage{times}
\usepackage{fullpage}
\usepackage[override]{cmtt}
\usepackage{sectsty}
\usepackage{graphicx}
\usepackage{color}
\usepackage{xspace}
\usepackage[lmargin=1in, rmargin=1in,tmargin=1in,bmargin=1in]{geometry}
\usepackage{url}
\usepackage{cite}
\usepackage{microtype}
\usepackage{listings}
\usepackage{subfig}

\newcommand{\TODO}[1]{\textcolor{red}{{\bf TODO:} #1}}
\newcommand{\ie}{{\em i.e.}}
\newcommand{\eg}{{\em e.g.}}
\newcommand{\benchmark}[1]{{\small\texttt{#1}}}
\newcommand{\ilcode}[1]{\texttt{#1}}
\newcommand{\comment}[1]{}

\newcommand{\gasnet}{GASNet}

\setlength{\parskip}{6pt}

\pagestyle{plain}

\begin{document}