### 2013-02-01 doc/asplos13/paper.tex

\input{header}
\usepackage{microtype}
\begin{document}

\title{Grappa: Scalable Computing on Commodity Hardware \\for Irregular Applications}
%\title{All Your Graphs Belong to Us}
%\authorinfo{\vspace{-2ex}Anonymous for review\vspace{-6ex}}{}{}
\authorinfo{Jacob Nelson$^{\dagger}$,
  Brandon Holt$^{\dagger}$,
  Brandon Myers$^{\dagger}$,
  Preston Briggs$^{\dagger}$,
  Luis Ceze$^{\dagger}$,
  Simon Kahan$^{{\dagger \ddagger}}$,
  Mark Oskin$^{\dagger}$
}{\textdagger University of Washington, \textdaggerdbl Pacific Northwest National Laboratory}{
  \{nelson, bholt, bdmyers, preston, luisceze, skahan, oskin\}@cs.washington.edu}

\maketitle
\begin{abstract}
\input{abstract}
\end{abstract}

\input{intro}

\input{relatedwork}

\input{overview}

\input{tasks}

\input{memory}

\input{communication}

\input{method}

\input{evaluation}

\input{conclude}

%\input{acknowledgements}

%\subsection{References}
\bibliographystyle{abbrv}
\bibliography{paper}

\end{document}