### 2013-02-12 doc/xps2013/header.tex
4l

\documentclass[11pt]{article}

\usepackage{times}
\usepackage{fullpage}
\usepackage[override]{cmtt}
\usepackage{sectsty}
\usepackage{graphicx}
\usepackage{color}
\usepackage{xspace}
\usepackage[lmargin=1in, rmargin=1in,tmargin=1in,bmargin=1in]{geometry}
\usepackage{url}
\usepackage{cite}
\usepackage{microtype}
\usepackage{listings}
\usepackage{subfig}

\newcommand{\TODO}[1]{\textcolor{red}{{\bf TODO:} #1}}
\newcommand{\ie}{{\em i.e.}}
\newcommand{\eg}{{\em e.g.}}
\newcommand{\benchmark}[1]{{\small\texttt{#1}}}
\newcommand{\ilcode}[1]{\texttt{#1}}
\newcommand{\comment}[1]{}

\newcommand{\gasnet}{GASNet}

\newcommand{\BigO}[1]{$O(#1)$}

\newcommand{\api}{\textls*{\small API}}
\newcommand{\apis}{\textls*{\small API}\hspace{0.75pt}s}
\newcommand{\ssd}{\textls*{\small SSD}}
\newcommand{\ssds}{\textls*{\small SSD}\hspace{0.75pt}s}

\setlength{\parskip}{6pt}

\pagestyle{plain}

\makeatletter
\renewcommand{\paragraph}{%
  \@startsection{paragraph}{4}%
  {\z@}{1.0ex \@plus 1ex \@minus .2ex}{-1em}%
  {\normalfont\normalsize\bfseries}%
}

\renewcommand\section{\@startsection{section}{1}{\z@}{-10\p@ \@plus -4\p@ \@minus -4\p@}{5\p@ \@plus 4\p@ \@minus 4\p@}{\normalfont\Large\bfseries\boldmath\rightskip=\z@ \@plus 8em\pretolerance=10000 }}

\renewcommand\subsection{\@startsection{subsection}{2}{\z@}{-10\p@ \@plus -4\p@ \@minus -4\p@}{5\p@ \@plus 4\p@ \@minus 4\p@}{\normalfont\large\bfseries\boldmath\rightskip=\z@ \@plus 8em\pretolerance=10000 }}

\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}{-10\p@ \@plus -4\p@ \@minus -4\p@}{5\p@ \@plus 4\p@ \@minus 4\p@}{\normalfont\normalsize\bfseries\boldmath\rightskip=\z@ \@plus 8em\pretolerance=10000 }}

\makeatother

\begin{document}