### 2013-03-23 doc/sosp13/abstract.tex
minor edits

Grappa is a runtime system for commodity clusters of multicore computers that
presents a massively parallel, single address space abstraction to
applications. Grappa's goal is to provide scalable performance for irregular
parallel applications, such as graph processing. These applications are
challenging because they exhibit little data locality, imbalanced parallel
work and complex communication patterns.

Grappa serves both as a C++ user library and as a foundation upon which higher
level languages can be developed or adapted. Grappa tolerates delays to remote
memory by multiplexing thousands of lightweight execution contexts to each
processor core; balances load via fine-grained distributed work-stealing;
takes full advantage of network characteristics by aggregating smaller data
requests into large ones; and provides efficient synchronization and remote
operations. We present a detailed description of the Grappa system,
programming examples using the library interface, and compare performance for
several irregular benchmarks to hand-optimized MPI code and the Cray XMT, a
custom system used to target the real time graph analytics market.