### 2013-03-27 doc/sosp13/abstract.tex
edits to abstract and intro

\Grappa is a runtime system for commodity clusters of multicore computers that
presents a massively parallel, single address space abstraction to
applications. \Grappa's purpose is to enable scalable performance of irregular
parallel applications, such as graph processing. Poor data locality,
imbalanced parallel work and complex communication patterns make scaling these
applications difficult.

\Grappa serves both as a C++ user library and as a foundation upon which
higher level languages can be developed or adapted. \Grappa tolerates delays
to remote memory by multiplexing thousands of lightweight workers to each
processor core; balances load via fine-grained distributed work-stealing;
increases communication throughput by aggregating smaller data requests into
large ones; and provides efficient synchronization and remote operations. We
present a detailed description of the \Grappa system and performance
comparisons on several irregular benchmarks to hand-optimized MPI code and to
the Cray XMT, a custom system used to target the real time graph analytics
market.