### 2013-03-28 doc/sosp13/relatedwork.tex

\section{Background}

\vspace{0.5ex}
\noindent{\bf Distributed graph processing} While \Grappa is a general runtime
system for any large-scale concurrent application, one of its killer
applications is graph analysis. Therefore, distributed graph processing
frameworks like Pregel~\cite{pregel:2010} and Distributed
GraphLab~\cite{distgraphlab:vldb12} are related. Pregel adopts a
bulk-synchronous parallel (BSP) execution model, which makes it inefficient on
workloads that could prioritize vertices. GraphLab overcomes this limitation
with an execution mode that schedules vertex computations individually,
allowing prioritization, which gives faster convergence in a variety of
iterative algorithms. GraphLab, however, imposes a rigid computation model
where programmers must express computation as transformations on a vertex and
its edge list only, with information only from adjacent vertexes. Pregel is
only slightly less restrictive, as the input data can be any vertex in the
graph. \Grappa also supports dynamic parallelism with asynchronous execution,
but parallelism is expressed as tasks or loop iterations, which is a far more
general programming model for irregular computation tasks.
PowerGraph~\cite{powergraph:osdi12} improves the performance of GraphLab for
real-world graphs with power-law degree distributions by using a vertex cut
for graph partitioning. If the size of the partial result of applying a subset
of edge data to the vertex is constant, then there is a good bound on the
communication required to keep vertex replicas coherent. We do not explore
these algorithmic techniques in this work, but they are complimentary and
would improve the performance of graph applications on \Grappa.

\vspace{0.5ex}
\noindent{\bf Software distributed shared memory.} Many traditional software
DSM systems are page based~\cite{Treadmarks,munin} and aim to hide the fact
that they are built in software from applications by exploiting the
processor's paging mechanisms, therefore relying heavily on locality. Instead,
\Grappa, like partitioned global address space (PGAS) models, implements its
DSM at the language, rather than the system level. Languages such as
Chapel~\cite{Chamberlain:2007}, X10~\cite{X10:2005}, and UPC~\cite{upc:2005}
make accesses to shared structures look like normal references. As we describe
later, \Grappa chooses a middle ground, where global addresses are explicit in
the API and local accesses are emitted conventionally by the compiler. Past
DSM work was mostly page-based, so it could directly take advantage of moving
large page-sized blocks of data from node to node. However, when commodity
networks move small blocks of data (a few bytes), only a fraction of the
available bandwidth is achieved. \Grappa had to include significant
aggregation support to make effective use of of the network.

In summary, while \Grappa's DSM system is conceptually similar to prior work,
we accept the random access aspect of irregular applications and optimize for
throughput rather than low latency. Our DSM system must support enough memory
concurrency to tolerate the latency of the network and additional latency
overhead impose by the runtime system.

\vspace{0.5ex}
\noindent{\bf Multithreading} \Grappa uses multithreading to tolerate memory
latency. This is a well known technique. Hardware implementations include the
Tera MTA~\cite{tera:mta1}, Cray XMT~\cite{feo:xmt}, Simultaneous
multithreading~\cite{tullsen:smt}, MIT Alewife~\cite{agarwal:alewife},
Cyclops~\cite{almasi:cyclops}, and even GPUs~\cite{gpus}. As we describe in
this paper, \Grappa implements its own software-based multithreading with a
lightweight user-mode task scheduler to multiplex \emph{thousands\/} of tasks
on a single processing core. The large number of tasks is required to tolerate
the very high inter-node communication latency of commodity networks.
\Grappa's task library employs several optimizations: an extremely fast task
switch, a small task size, and judicious use of software prefetch instructions
to bring task state into the cache sufficiently long before that task is
actually scheduled. The main difference between \Grappa's support for
lightweight threads and prior work such as
QThreads~\cite{Wheeler08qthreads:an} and
Capriccio~\cite{Behren03capriccio:scalable} is context prefetching, which is needed for good performance when multiplexing such a large
number of tasks.