### 2013-03-28 doc/sosp13/method.tex

\section{Methodology} \label{sec:method}

We implemented the \Grappa in C++ for the Linux operating system. The core
runtime system system is about 15K lines of code. We ported a number of
benchmarks to \Grappa as well as collected and optimized a set of comparison
benchmarks for XMT, MPI and UPC~\cite{UPC}. We run the \Grappa, MPI and UPC
experiments on a cluster of AMD Interlagos processors. Nodes have 32 2.1-GHz
cores in two sockets, 64GB of memory, and 40Gb Mellanox ConnectX-2 InfiniBand
network cards. Nodes are connected via a QLogic InfiniBand switch. We also
compare \Grappa to a 128-node Cray XMT (3rd generation MTA). Each node
consists of a 500-MHz MTA Threadstorm multithreaded processor that supports
128~streams. The machine uses Cray's proprietary SeaStar2 interconnection
network.

We use a variety of benchmarks:

\noindent{\bf Context-switch:} This simple microbenchmark is used to explore context switch overhead.  A configurable number of workers are used to increment the values in a large array.  This benchmark stresses the scheduler inside of Grappa.

\nonindent{\bf GUPS:} This is a faithful implementation of the
giga-updates-per-second benchmark. Read-modify-write updates are dispatched at
random to a large array. This benchmark stresses the networking layer of
Grappa separately from the scheduler, because only a single worker is used per
system node.

\nonindent{\bf Unbalanced tree search in-memory (UTS-Mem):} Unbalanced Tree Search
(UTS) is a benchmark for evaluating the programmability and performance of
systems for parallel applications that require dynamic load
balancing~\cite{UTS}. It involves traversing an unbalanced implicit tree: at
each vertex, its number of children is sampled from some probability
distribution, and this number of new nodes are added to a work queue to be
visited. While this benchmark captures irregular, dynamic \emph{computation,}
we actually want to evaluate performance of algorithms with irregular
\emph{memory\/} access patterns. Thus we augment UTS by using the existing
traversal code to create a large tree in memory, and then we traverse the
in-memory tree. We call this benchmark UTS-Mem, and the timed portion is this
traversal of the in-memory tree. This in-memory traversal has no knowledge of
the tree structure beforehand.

\nonindent{\bf Breadth-first-search (BFS):} This is the primary kernel for the
Graph500 benchmark and is what currently determines the ranking of machines on
the Graph500 list~\cite{graph500list}. As a whole, the Graph500 benchmark
suite is designed to bring the focus of system design on data-intensive
workloads, particularly large-scale graph analysis problems, that are
important among cybersecurity, informatics, and network understanding
workloads. The BFS benchmark builds a search tree containing parent nodes for
each traversed vertex during the search. While this is a relatively simple
problem to solve, it exercises the random-access and fine-grained
synchronization capabilities of a system as well as being a primitive in many
other graph algorithms. Performance is measured in \emph{traversed edges per
second\/} (TEPS), where the number of edges is the edges making up the
generated BFS tree. One of the reference implementations of Graph500 BFS is
for the XMT; this code fails to scale past 16 XMT processors because it does
not expose enough parallelism, so we modified the code to use a recursive loop
decomposition similar to Grappa's. We compare this modified version against a
straightforward Grappa implementation. We do not employ algorithmic
improvements, though there are
many~\cite{Beamer:Graph500,Yoo:FixedPointGraph500}.

\nonindent{\bf IntSort:} \TODO{Write me Brandon Holt!}

\nonindent{\bf PageRank:} \TODO{Write me Brandon Myers!}

The metric we use is algorithmic time, which means startup and loading of the
data structure (from disk) is not included in the measurement. \Grappa
collects statistics about application behavior (packets sent, context
switches, etc) and these are discussed where appropriate. In addition we use a
profiled version of \Grappa to report component runtimes, but this profiled
version is only used to collect this data and a fully optimized version is
used for the bulk of the presented results.

%%%% Bit bucket

%\paragraph{Betweenness Centrality} An important measure of the importance of particular vertices in a network is betweenness centrality (BC)~\cite{freeman1979centrality}. By this measure, the ``importance'' of each vertex is computed by finding the fraction of shortest paths that pass through it, which can optionally be approximated by only computing shortest paths using a subset of the vertices as starting points. BC can be useful for understanding which vertices may have the greatest impact, so in social networks this could be the primary person linking two communities. Because it requires multiple breadth-first traversals across the entire graph and in reverse, on power-law degree graphs, this algorithm exercises random accesses rate, and load balancing. It also requires fine-grained synchronization on updates to vertex centrality values because multiple paths will update the same vertex. BC is a kernel in the DARPA High Performance Computing Systems (HPCS) Scalable Synthetic Compact Applications graph analysis (SSCA\#2) benchmark~\cite{ssca2}. Performance for BC is also measured in TEPS, where the number of traversed edges is the total number of edges in the graph multiplied by the number of random starting vertices used in the approximate computation. We use the XMT implementation of BC implemented as part of the GraphCT~\cite{GraphCT} library and a comparable Grappa implementation, both of which use the parallel BC algorithm developed by Bader et al.~\cite{bader:bc}.

%To explore the performance of the Grappa runtime we have implemented three algorithms: breadth first search, betweenness centrality, and unbalanced tree search.  These algorithms were implemented for the Cray XMT (our baseline) and for Grappa.  The metric we use is algorithmic time, which means startup and loading of the data structure (from disk) is not included in the measurement.  Data is collected on real systems, which means minor variations in runtime exist from run to run.  The average of multiple runs are used, and where appropriate, confidence in the result is reported.

\comment{The Grappa version of tree search UTS-mem

\lstset{language=C++,
       basicstyle=\footnotesize,
       tabsize=2}
\begin{lstlisting}
// base of shared vertex_t[]
GlobalAddress<vertex_t> Vertices;

// base of shared int64_t[]
GlobalAddress<int64_t> ChildrenPointers;

void search_vertex( id ) {
  GlobalAddress<vertex_t> v_addr = Vertices + id;
  Incoherent<vertex_t>::RO v( v_add, 1 );

  // start index of my children pointers
  childIndex = v->childIndex;

  // how many children I have
  numChildren = v->numChildren;

  // parallel loop over the child list for this vertex
  parallel_for(fn=&search_children, start=childIndex, iters=numChildren);
}

void search_children( start, iters ) {
  // take advantage of spatial locality in the array of children
  GlobalAddress<int64_t> child_base_addr = ChildrenPointers + start;
  Incoherent<int64_t>::RO childIds( child_base_addr, iters );

  // spawn a task to visit each child
  for (i = 0..iters) {
    SoftXMT_publicTask( fn=&search_vertex, id=childIds[i] );
  }
}
\end{lstlisting}
}

%% Given that we have MPI performance on the cluster this is not
%% that important to have here anymore.
%%
%%
%One question that must be answered when making a comparison between the Cray XMT and a typical HPC cluster is what is a fair comparison -- these systems are quite different.  Three options immediately come to mind: equal number of processing cores, equal number of network interfaces, and equal dollars.  We discounted the last option fairly quickly, because it isn't a lasting data-point -- the cost of hardware shifts over time, skewing the interpretation of results.  The first option, cores, has some merit, but Grappa is designed for applications that have no locality in their computation.  This means almost all of their memory accesses are remote.  The factor that limits their performance is not processing, but communicating.  Hence, we have chosen the middle option, network interfaces as the way to normalize across the XMT and our cluster.  Each processor in the XMT system has its own network interface to access shared memory.  In the HPC cluster we use, each processor (which contains up to 32~cores, although we only use 6 in our experiments), has a single infiniband interface.  Hence, for our results we scale up XMT processors one for one with full system nodes.

%Towards evaluating how well Grappa enables scalable performance of irregular applications on mass market systems, our methodology is as follows.  For consistency in measurement, all results reported for mass market systems are measured on the same platform.  To establish the potential for Grappa performance, we identify small microkernel benchmarks that provide insight into the limitations of our system's component performance behaviors.  To compare meaningfully with the state-of-the-art, we compare Grappa performance on well known irregular application benchmarks both against the popular high performance programming methodologies MPI and UPC for distributed memory clusters, and against a custom shared memory architecture used exclusively for irregular computation, the Cray XMT. We then examine how sensitive Grappa results are to reduced injection rate and increased latency, as often occur when a system is scaled up so that nodes are further apart and the effective per destination packet injection rate drops. The three sections that follow describe this methodology in greater detail.