### 2013-05-27 doc/pgas13/abstract.tex
Work on flat-combining section (also fix \TODO, add fc papers to bib)

TBD

% Structured concurrency (scheduler, cooperative multithreading + communication) enables hierarchical flat-combining to be implemented efficiently for a software PGAS shared-memory solution.