### 2013-07-19 doc/asplos14/asplos14-template/template.tex
Use ASPLOS14 style

\documentclass[pageno]{jpaper}

%replace XXX with the submission number you are given from the ASPLOS submission site.
\newcommand{\asplossubmissionnumber}{XXX}

\usepackage[normalem]{ulem}

\begin{document}

\title{
Instructions for Submissions to ASPLOS 2014}

\date{}
\maketitle

\thispagestyle{empty}

\begin{abstract}

This document provides instructions for submitting papers to the 19th
International Conference on Architectural Support for Programming Languages
and Operating Systems (ASPLOS), 2014.  It also provides a sample for how
papers must be formatted.
\end{abstract}

\section{Introduction}

This document provides instructions for submitting papers to the 19th
International Conference on Architectural Support for Programming Languages
and Operating Systems (ASPLOS), 2014.  Submissions that violate these
instructions may not be reviewed, at the discretion of the program chair.

An example submission (formatted using the ASPLOS'14 submission format)
that contains the submission and formatting guidelines can be downloaded
from here:
\href{http://www.cs.utah.edu/asplos14/files/asplos14-template.pdf}{Sample
PDF}. The contents of this document are the same as the contents of the
submission instructions that appear on
\href{http://www.cs.utah.edu/asplos14/submission.html}{this website}.  The
paper submission site is \href{http://asplos14.cs.illinois.edu}{here.}

All questions regarding paper formatting and submission should be directed
to the program chair.

\section{What's New?}

We specifically call attention to the following instructions/features new
to ASPLOS this year.

\begin{itemize} 
\item Text must be in 10pt ({\bf not} 9pt) font.
\item Papers must be at most 12 pages, not including references and an
optional appendix. 
\item Each reference must specify {\em all} authors (no {\em et al.}).
\item Authors may optionally suggest reviewers. 
\item Authors may optionally provide information on previous rejections.
\item Authors of {\em all} accepted papers will be required to give a
lightning presentation (about 90s) and a poster in addition to the regular
conference talk.
\item Proceedings will appear in the ACM digital library up to two weeks
before the conference.
\end{itemize}

\section{Acceptance Philosophy and Authors' Commitments}

Every effort will be made to judge a paper on its own merits. There will be
no target acceptance rate. We expect to accept a wide range of papers with
appropriate expectations for evaluation -- while papers that build on
significant past work with strong evaluations are valuable, papers that
open new areas with less rigorous evaluation are even more so.  The mantra
for this ASPLOS is: {\bf A}lways {\bf S}upport {\bf P}otentially {\bf
L}ong-term and {\bf O}ut-of-the-box {\bf S}ystems.

Accepting all worthy papers sometimes results in multitrack sessions,
reduced talk lengths, and reduced interaction. To mitigate these downsides
and in keeping with the spirit of ASPLOS as a melting pot for multiple
disciplines, we will have the following (borrowed from Micro'12).  We will
kick off the conference with a {\bf lightning presentation} session with a
(roughly) 90s presentation of each accepted paper. There will also be a {\bf
poster/demo session} where {\em all} accepted papers will be required to
participate with a poster and/or demo to facilitate closer interactions. Of
course, there will also be the regular conference talks for each paper.
Authors should be prepared to participate in all three sessions (lightning
presentation, poster/demo, and regular talk).

\section{The Paper}

\subsection{Formatting}

Submitted manuscripts should contain a maximum of 12 pages of single-spaced
two-column text, {\bf not including references and an optional
appendix}. You may include any number of pages for references and the
appendix, but see below for more instructions.  If you are using
\LaTeX~\cite{lamport94} to typeset your paper, then we suggest that you use
the template here:
\href{http://www.cs.utah.edu/asplos14/files/asplos14-latex-template.tar.gz}{\LaTeX~Template}.
(\href{http://www.cs.utah.edu/asplos14/files/asplos14-template.pdf}{This
document} was prepared with that template.)  If you are using a different
software package to typeset your paper, then please adhere to the
guidelines given in Table~\ref{table:formatting}.\footnote{One exception is
that authors may use the SIGPLAN style/class file
\href{http://classic.sigplan.org/sigplanconf.cls}{here}, but {\bf only with
the 10pt body font option (9pt will be rejected)} and modified as needed
for the requirements of the references section below.  This is marginally
different from the specified template, but will be accepted due to its
widespread use.}

\begin{table}[h!]
  \centering
  \begin{tabular}{|l|l|}
    \hline
    \textbf{Field} & \textbf{Value}\\
    \hline
    \hline
    Page limit & 12 pages, {\bf not including}\\
               & {\bf references \& appendix}\\
    \hline
    Paper size & US Letter 8.5in $\times$ 11in\\
    \hline
    Top margin & 1in\\
    \hline
    Bottom margin & 1in\\
    \hline
    Left margin & 0.75in\\
    \hline
    Right margin & 0.75in\\
    \hline
    Separation between columns & 0.25in\\
    \hline
    Body font & 10pt\\
    \hline
    Abstract font & 10pt, italicized\\
    \hline
    Section heading font & 12pt, bold\\
    \hline
    Subsection heading font & 10pt, bold\\
    \hline
    Caption font & 9pt, bold\\
    \hline
    References & 8pt\\
    \hline
  \end{tabular}
  \caption{Formatting guidelines for submission. }
  \label{table:formatting}
\end{table}

\textbf{Please ensure that you include page numbers with your
submission}. This makes it easier for the reviewers to refer to different
parts of your paper when they provide comments.

Please ensure that your submission has a banner at the top of the title
page, similar to
\href{http://www.cs.utah.edu/asplos14/files/asplos14-template.pdf}{this
one}, which contains the submission number and the notice of
confidentiality.  If using the template, just replace XXX with your
submission number.

\subsection{Content}

\noindent\textbf{\sout{Author List.}} Reviewing will be double blind;
therefore, please do not include any author names on any submitted
documents except in the space provided on the submission form.  You must
also ensure that the metadata included in the PDF does not give away the
authors. If you are improving upon your prior work, refer to your prior
work in the third person and include a full citation for the work in the
bibliography.  For example, if you are building on {\em your own} prior
work in the papers \cite{nicepaper1,nicepaper2,nicepaper3}, you would say
something like: "While the authors of
\cite{nicepaper1,nicepaper2,nicepaper3} did X, Y, and Z, this paper
additionally does W, and is therefore much better."  Do NOT omit or
anonymize references for blind review.  There is one exception to this for
your own prior work that appeared in IEEE CAL, workshops without archived
proceedings, etc.\, as discussed later in this document.

\noindent\textbf{Figures and Tables.} Ensure that the figures and tables
are legible.  Please also ensure that you refer to your figures in the main
text.  Many reviewers print the papers in gray-scale. Therefore, if you use
colors for your figures, ensure that the different colors are highly
distinguishable in gray-scale.

\noindent\textbf{References.}  There is no length limit for references.
{\bf Each reference must explicitly list all authors of the paper.  Papers
not meeting this requirement will be rejected.} Authors of NSF proposals
should be familiar with this requirement. Knowing all authors of related
work will help find the best reviewers.

\noindent\textbf{Optional Appendix.}  Authors may optionally include an
appendix (no length limit) as the last section of the manuscript; however,
{\em reviewers are not obliged to read the appendix.}  An appendix may
include proofs of theorems, more details on methodology, more results, and
anything else that can potentially answer reviewer questions.  The rest of
the manuscript may cite the appendix, but the paper should stand on its own
without the appendix.  Authors need not feel compelled to include an
appendix -- we understand the authors' time is best spent on the main
manuscript.

\section{Other Information for the Submission Site}

\subsection{Declaring Authors}

Declare all the authors of the paper upfront. Addition/removal of authors
once the paper is accepted will have to be approved by the program chair,
since it potentially undermines the goal of eliminating conflicts for
reviewer assignment.

\subsection{Areas and Topics}

ASPLOS emphasizes multidisciplinary research. Submissions should ideally
emphasize synergy of two or more ASPLOS areas: architecture, programming
languages, operating systems, and related areas (broadly
interpreted). Authors should indicate these areas on the submission form as
well as specific topics covered by the paper for optimal reviewer match. If
you are unsure whether your paper falls within the scope of ASPLOS, please
check with the program chair -- ASPLOS is a broad, multidisciplinary
conference and encourages new topics.

\subsection{Declaring Conflicts of Interest}

Authors must register all their conflicts on the paper submission site.
Conflicts are needed to ensure appropriate assignment of reviewers.  
If a paper is found to have an undeclared conflict that causes
a problem OR if a paper is found to declare false conflicts in order to
abuse or ``game'' the review system, the paper may be rejected.

Please declare a conflict of interest (COI) with the following 
for any author of your paper:

\begin{enumerate}
\item Your Ph.D. advisor, post-doctoral advisor, and Ph.D. students.
\item Family relations by blood or marriage (if they might be potential reviewers).
\item People with whom you co-authored accepted/rejected/pending papers in the last five years.
\item People with whom you co-authored accepted/pending grant proposals
  in the last five years.
\item People who shared your primary institution in the last five years. 
\end{enumerate}

``Service'' collaborations like co-authoring a CSTB report or co-presenting
tutorials do not constitute conflicts.  Co-authoring a paper that is a
compendium of various projects with no true collaboration among the
projects does not constitute a conflict among the authors of the different
projects.  On the other hand, there may be others not covered by the above
with whom you know a COI exists.  Please report such COIs; however, you
will need to justify them. Please be reasonable.  For example, just because
a reviewer works on similar topics as your paper, you cannot declare a COI
with that reviewer. We will carefully check the justification of conflicts.

We hope to draw most reviewers from the PC and the ERC, but others from the
community may also write reviews.  Please declare all your conflicts (not
just restricted to the PC and ERC).  When in doubt, contact the program
chair.

\subsection{Optional Information on Previous Submissions}

In addition to your paper, you may optionally upload your past similar
(rejected) submissions, their reviews, the author responses, and/or a
statement of how your ASPLOS'14 submission improved over the previous
submissions or addressed the previous reviews. Please use only one PDF file
for all this information and include the venue(s) where the previous work
was submitted.  Please do not include any material that will compromise
author anonymity.

Each paper will undergo a full new review process independent of whether
such a PDF file is submitted. The ASPLOS'14 reviewers will see this
document (and learn about its existence) only after they have finalized
their review.  It is, however, possible that reading this document will
influence the ASPLOS reviewers to request a change of their review or
request the PC chair to get another opinion from one of the previous
reviewers. Please consider all such implications carefully before
submitting the PDF file.

Authors are under no obligation to submit such a document. Reviewers
will be given explicit instructions not to hold the lack of such a document
against the paper in any way. Similarly, reviewers are under no obligation
to read such a document (unless a review brings up a past
submission explicitly).

\subsection{Optional Reviewer Suggestions}

Authors may optionally mark (non-conflicted) PC and ERC members that they
believe could provide expert reviews for their submission.  If authors
believe there is insufficient expertise on the PC and ERC for the topic of
their paper, they may suggest alternate reviewers.  The program chair will
use the authors' input at her discretion.  We provide this opportunity
for input mostly for papers on non-traditional and emerging topics.

\subsection{Concurrent Submissions and Workshops}

By submitting a manuscript to ASPLOS'14, the authors guarantee that the
manuscript has not been previously published or accepted for publication in
a substantially similar form in any conference, journal, or the archived
proceedings of a workshop (e.g., in the ACM digital library) -- see
exceptions below. The authors also guarantee that no paper that contains
significant overlap with the contributions of the submitted paper will be
under review for any other conference or journal or an archived proceedings
of a workshop during the ASPLOS'14 review period. Violation of any of these
conditions will lead to rejection.

The only exceptions to the above rules are for the authors' own papers in
(1) workshops without archived proceedings such as in the ACM digital
library (or where the authors chose not to have their paper appear in the
archived proceedings), or (2) venues such as IEEE CAL where there is an
explicit policy that such publication does not preclude longer conference
submissions.  In all such cases, the submitted manuscript may ignore the
above work to preserve author anonymity. This information must, however, be
provided on the submission form -- the PC chair will make this information
available to reviewers if it becomes necessary to ensure a fair review.
(This policy will be explicitly conveyed to the reviewers.)

As always, if you are in doubt, it is best to contact the program chair.

Finally, we also note that the ACM Plagiarism Policy ({\em
http://www.acm.org/publications/policies/plagiarism\_policy}) covers a
range of ethical issues concerning the misrepresentation of other works or
one's own work.

\section{Early Access in the Digital Library}

The ASPLOS'14 proceedings will be freely available via the ACM Digital
Library for up to two weeks before and up to a month after the
conference. {\bf Authors must consider any implications of this early
disclosure of their work {\em before} submitting their papers.}

\section{Acknowledgements}
Several ideas in this document and parts of the document text have been
taken from previous conferences and we thank their program chairs;
especially, Christos Kozyrakis (Micro'13), Margaret Martonosi (ISCA'13),
and Onur Mutlu (Micro'12).

\bibliographystyle{plain}
\bibliography{references}

\end{document}