### 2013-07-22 doc/asplos14/overview.tex

\section{\Grappa Overview}

Commodity clusters provide fast processors and high-bandwidth
interconnects but suffer when they encounter unpredictable data access
patterns. Maximizing throughput in such a situation depends upon
proper management of concurrency and communication. \Grappa's design is
motivated by four relationsips we observe within commodity clusters.

First, cores in our cluster can execte hundreds of thousands of
instructions in the few microseconds that it takes for one RDMA read
to complete. Given the high degree of fine-grained random access in
irregular applications, this pushes us to overlap communication with
other computation.

Second, cores can perfom random access to local DRAM over two orders
of magnitude faster than they can issue RDMA requests through the
network. In addition, cores can execute instructions three orders of
magnitude faster than they can issue RDMA requests. On the other hand,
per-node network bandwidth (measured in bytes rather than messages) is
within an order of magnitude of local DRAM bandwidth. The combination
of these facts motivates us to delay and aggregate remote operations
to run on remote cores.

As \Grappa depends on concurrency for performance, its programming
model focuses on enabling programmers to express large amounts of
concurrency. \Grappa is intended to support what we think is a
desriable workflow: programmers start by expressing a computation
using a multithreaded, shared-memory model on a single node. \Grappa
then provides the primitves to do a straightforward translation of
this code to run on a distributed-memory cluster, through constructs
for optimizing locality and overlapping communication.

\Grappa comprises three main software components, shown in Figure~\ref{fig:grappa}:

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.95\columnwidth]{figs/system-overview}
\begin{minipage}{0.95\columnwidth}
  \caption{\label{fig:grappa} \Grappa system overview}
\end{minipage}
\vspace{-3ex}
\end{center}
\end{figure}

\begin{description}

\item [Tasking system.] The tasking system supports lightweight multithreading
to tolerate communication latency and global distributed work-stealing (i.e.,
tasks can be stolen from any node in the system), which provides automated
load balancing. The scheduler oversubscribes to have more worker threads than
required for latency tolerance. By keeping at least four workers per 
core ready to run at all times, the scheduler can prefetch a worker's state
into cache to reduce the chance of stalling on memory accesses during a
context switch.

\item[Communication layer.] The main goal of our communication layer is to
aggregate small messages into large ones.  This process is invisible to the application
programmer. Its interface is based on active messages~\cite{vonEicken92}.
Since aggregation and deaggregation of messages needs to be very efficient, we perform the process in parallel and carefully use lock-free synchronization
operations.
\TODO{Re-include the mention of GASNET,MPI spawning,and inifiniband
    somewhere but not in the overview intro}

\item[Distributed shared memory.] The DSM system provides fine-grain access to
data anywhere in the system, with delegated operation at the core of its design.
Every piece of global memory is owned by a particular core in the system, and all others may only access that memory by delegating their requests to the owning core.
It supports normal access operations such as
\emph{read\/} and \emph{write\/} as well as synchronizing operations such as
\emph{fetch-and-add\/}~\cite{fetchandadd}. Due to delegation, the memory model offered is similar to what underpins
C/C++~\cite{N2480,N2800}, so it is familiar to programmers. The DSM system
design relies on the lightweight tasking system and communication layer in
order to offer high aggregate random access bandwidth for accessing remote
data.

\end{description}