### 2013-07-23 doc/asplos14/conclude.tex
reworded confusing sentance

\section{Conclusion}

Irregular computations are both important and challenging to execute
quickly.  Scaling these applications easily on commodity hardware has
been a historical challenge. \Grappa is a runtime framework that
simplifies this task for software developers and compiler
writers. This paper describes the \Grappa framework and its three main
components: a task library, a distributed shared memory system, and a
network aggregator. \Grappa's key aspect is extreme latency tolerance,
which not only hides network latency but also enables the system to
spend time on sophisticated work stealing and network optimizations,
trading latency for even more throughput.

Our evaluation of \Grappa reveals that the core components, scheduling
and communication, achieve their design goals.  Thousands of workers
can be efficiently context switched on a multicore processor, up to
the DRAM bandwidth.  Aggregating messages enables \Grappa to achieve
over 1.0~GUPS on 64~nodes.  We also explored four other algorithms:
unbalanced tree search, breadth first search, PageRank, and integer
sort, comparing performance of these algorithms on \Grappa, the Cray
XMT, and hand-optimized MPI implementations.  As might be expected,
when an MPI implementation exploits, in the application code, many of the
same techniques \Grappa provides in the runtime, such as aggregating
requests, yet without the overhead of tasking, the MPI implementation
excels.  This comes with enormous implementation complexity, burdening
the application developer. Moreover, significant tuning has yet to be
done on the \Grappa runtime system.  Compared to the Cray XMT, \Grappa
is between 2.6X faster and 4.3X slower.  Yet when cost-performance is
considered, mass-market x86-based cluster hardware makes \Grappa a
highly attractive option.