### 2013-07-24 doc/asplos14/conclude.tex

\section{Conclusion}

Irregular computations are both important and challenging to execute
quickly.  Scaling these applications easily on commodity hardware has
been a historical challenge. \Grappa 
simplifies this task for software developers and compiler
writers. 
% This paper describes \Grappa's framework and its main
% components: a task library, a distributed shared memory system, and a
% network aggregator. 
\Grappa's key aspect is extreme latency tolerance,
which not only hides network latency but also enables the system to
spend time on sophisticated work stealing and network optimizations,
trading latency for even more throughput.

Our evaluation of \Grappa shows that the core components, scheduling
and communication, achieve their design goals.  Thousands of workers
can be efficiently context switched on a multicore processor, limited by DRAM bandwidth.  Aggregating messages enables \Grappa to achieve
over 1.0~GUPS on 64~nodes.  We also explored four other algorithms:
unbalanced tree search, breadth first search, PageRank, and integer
sort, comparing performance of these algorithms on \Grappa, the Cray
XMT, and optimized MPI implementations. \Grappa is between 11X faster and 5.4X
slower than MPI and UPC, and between 2.6X faster and 4.3X slower than
the XMT. When \Grappa is slower than MPI, we find that the MPI
implementation includes a specialized implementation of a feature
\Grappa provides generally to all applications. This generality comes
at a performance cost but eases the application developer's
task. Moreover, there are significant tuning opportunities the \Grappa
runtime system.  When cost-performance is considered, \Grappa on mass-market
x86-based clusters is a highly attractive option.