### 2013-09-10 doc/pgas13/conclusion.tex

\section{Conclusion}
% At the core of this work is the desire to have a library of data structures which can be used without worrying about when they will be consistent or whether they will limit scalability.
% Bringing findings from shared memory into the PGAS domain brings with it new challenges and opportunities.
Coming from the multi-core domain, the flat-combining paradigm provides a new perspective to global synchronization, bringing with it both challenges and new opportunities.
We have shown that the additional concurrency that comes with a latency-tolerant runtime, rather than compounding the problem, provides a new opportunity for reducing communication by combining locally.
% perform some of the locality optimizations that finely tuned implementations would.
In PGAS implementations there is typically a large discrepancy between the first simple description of an algorithm and the final optimized one.
Our distributed flat-combining framework allows easy implementation of a library of flat-combined linearizable global data structures, allowing even simple applications that use them to scale out to a thousand cores and millions of concurrent threads.